Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-16 csf-2017/paper.tex
Fix the sentence that Lau noticed was broken.
LR changes

For technical reasons, we do not actually remove a revoked temporary region from
the world, but we turn it a special revoked region that exists for this purpose.
Such a revoked region contains no state transition system and puts no
requirements on the memory. It simply serves as a mask for a revoked temporary
region. Masking a region like this goes back to earlier work of
\citet{Ahmed2004semantics} and was also used by
\citet{Thamsborg:2011:KLR:2034773.2034831}.

@@

The \emph{public future} region relation is the simplest. It satisfies
% \dominique{is defined by?} 
% \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations. It can be shown that these will have the properties mentioned below}
the following properties:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
    (s,s') \in \phi_\pub }
  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
Both temporary and permanent regions are only allowed to transition according to
the public part of their transition system. Additionally, revoked regions must
either remain revoked or be replaced by a temporary region. This means that the
public future world relations allows us to reuse a region that has been revoked
earlier.

The \emph{private future region} relation satisfies:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
              (s,s') \in \phi} 
  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr \revoked }
\end{mathpar}
Here, revocation of temporary regions is allowed. As explained above, this can
be done by replacing them with a special revoked region, which does not describe
any memory, but in fact, any other region can also be used to replace both
revoked and temporary regions. On the other hand, permanent regions cannot be
masked away. They are only allowed to transition according to the private part
of the transition system.

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{
    \begin{multlined}
      \npairP{(\var{perms}, \\
        \start,\addrend)}
    \end{multlined}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
}
%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation now defines semantically when values, terms, and
configurations are for execution by the adversary, i.e. without the risk that
they will be able to use it to break protocols in the world. The definition of
the logical relation is found in \figurename{}s~\ref{fig:logrel}
and~\ref{fig:perm-conds} and we provide some explanations in the following
paragraphs. We omit some definitions and explain them only verbally, but precise
definitions can be found in the technical appendix~\cite{technical_appendix}.

@@

In order to formulate results about \texttt{\footnotesize{f1}} and
\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
fails. To this end, we assume they have access to a flag (an address in memory)
that they set to $1$ if the assertion fails and then halt. The correctness lemma
for \texttt{\footnotesize{f1}} then states:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
where each of the capabilities have an appropriate range of authority and
pointer\footnote{These assumptions are kept intentionally vague for brevity.
  Full statements are in the technical appendix~\citep{technical_appendix}.}.
Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

@@

%\section{Conclusion}
%\label{sec:conclusion}
\todo[inline]{Insert the correct URL for the technical report in the bibtex file!}

>>>>

For technical reasons, we do not actually remove a revoked temporary region from
the world, but we turn it into a special revoked region that exists for this purpose.
Such a revoked region contains no state transition system and puts no
requirements on the memory. It simply serves as a mask for a revoked temporary
region. Masking a region like this goes back to earlier work of
\citet{Ahmed2004semantics} and was also used by
\citet{Thamsborg:2011:KLR:2034773.2034831}.

@@

The \emph{public future} region relation is the simplest. It satisfies
% \dominique{is defined by?} 
% \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations. It can be shown that these will have the properties mentioned below}
the following properties:
\begin{mathpar}
  \inferrule{ (s,s') \in \phi_\pub }
  {  (v,s',\phi_\pub,\phi,H) \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
Both temporary and permanent regions are only allowed to transition according to
the public part of their transition system. Additionally, revoked regions must
either remain revoked or be replaced by a temporary region. This means that the
public future world relations allows us to reuse a region that has been revoked
earlier.

The \emph{private future region} relation satisfies:
\begin{mathpar}
  \inferrule{ (s,s') \in \phi} 
  {  (v,s',\phi_\pub,\phi,H) \futurestr (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr \revoked }
\end{mathpar}
Here, revocation of temporary regions is allowed. As explained above, this can
be done by replacing them with a special revoked region, which does not describe
any memory, but in fact, any other region can also be used to replace both
revoked and temporary regions. On the other hand, permanent regions cannot be
masked away. They are only allowed to transition according to the private part
of the transition system.

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{
    \begin{multlined}
      \npairP{(\var{perms}, \\
        \start,\addrend)}
    \end{multlined}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
}
%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation now defines semantically when values, terms, and
configurations can safely be used or executed by the adversary, i.e. without the
risk that they will be able to use it to break protocols in the world. The
definition of the logical relation is found in \figurename{}s~\ref{fig:logrel}
and~\ref{fig:perm-conds} and we provide some explanations in the following
paragraphs. We omit some definitions and explain them only verbally, but precise
definitions can be found in the technical appendix~\cite{technical_appendix}.

@@

In order to formulate results about \texttt{\footnotesize{f1}} and
\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
fails. To this end, we assume they have access to a flag (an address in memory)
that they set to $1$ if the assertion fails and then halt. The correctness lemma
for \texttt{\footnotesize{f1}} then states:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
where each of the capabilities have an appropriate range of authority and
pointer\footnote{These assumptions are kept intentionally vague for brevity.
  Full statements are in the technical appendix~\citep{technical_appendix}.}.
Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

@@

%\section{Conclusion}
%\label{sec:conclusion}
